---
title: "Law Ai Medicine"
output: html_notebook
---

```{r}
library(hrbrthemes)
library(viridis)
library(plyr)
library(ggplot2)
library(grid)
library(knitr)
library(RColorBrewer)
library(tidyverse)
library(countrycode)
```

```{r, echo= FALSE }
#read new data
dt = read.csv("Book1.csv",skip = 1)
dt2 = read.csv("Book2.csv",skip = 1)

#merge together on item number
data = merge(dt,dt2,by = "Item.Number")
```

## Bubble plots and plots broken down by year

```{r , echo= FALSE }
# Transform data to count Field of medicine and legal issue Mentioned by year
yr = c(1)
field = c(1)
Men = c(1)
new_df = data.frame(yr,field,Men)

for(i in 1:nrow(data)) {
    row <- data[i,]
    # create row if an article mentions Privacy
    if(row$Privacy.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      fld = row$Field.of.Medicine.1
      Mention  = "Privacy"
      new_df <- rbind(new_df,c(yr,fld, Mention))
      # datalist[[i]] = c(yr,fld,Mention)
    }
    
    # create row if an article mentions Cyber Security
    if(row$Cybersecurity.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      fld = row$Field.of.Medicine.1
      Mention  = "Cyber Security"
      new_df <- rbind(new_df,c(yr,fld, Mention))
    }

    # create row if an article mentions Safety/Quality
    if(row$Safety.Quality.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      fld = row$Field.of.Medicine.1
      Mention  = "Safety/Quality"
      new_df <- rbind(new_df,c(yr,fld, Mention))
    }

    # create row if an article mentions Bias
    if(row$Bias.Discrimination.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      fld = row$Field.of.Medicine.1
      Mention  = "Bias/Discrimination"
      new_df <- rbind(new_df,c(yr,fld, Mention))
    }

    # create row if an article mentions Liability
    if(row$Liability.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      fld = row$Field.of.Medicine.1
      Mention  = "Liability"
      new_df <- rbind(new_df,c(yr,fld, Mention))
    }

    # create row if an article mentions Informed Consent
    if(row$Informed.Consent.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      fld = row$Field.of.Medicine.1
      Mention  = "Informed Consent"
      new_df <- rbind(new_df,c(yr,fld, Mention))
    }

    # create row if an article mentions Capacity
    if(row$Capacity.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      fld = row$Field.of.Medicine.1
      Mention  = "Capacity"
      new_df <- rbind(new_df,c(yr,fld, Mention))
    }

    # create row if an article mentions IP
    if(row$Intellectual.Property.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      fld = row$Field.of.Medicine.1
      Mention  = "Intellectual Property"
      new_df <- rbind(new_df,c(yr,fld, Mention))
    }

    # create row if an article mentions EOR
    if(row$Efficiency.of.Regulation.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      fld = row$Field.of.Medicine.1
      Mention  = "Efficiency of Regulation"
      new_df <- rbind(new_df,c(yr,fld, Mention))
    }

    # create row if an article mentions AI and legal
    if(row$AI.Personhood.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      fld = row$Field.of.Medicine.1
      Mention  = "AI and Legal Personhood"
      new_df <- rbind(new_df,c(yr,fld, Mention))
    }

    # create row if an article mentions Access/Insurance
    if(row$Access.Insurance.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      fld = row$Field.of.Medicine.1
      Mention  = "Access/Insurance"
      new_df <- rbind(new_df,c(yr,fld, Mention))
    }


    # create row if an article mentions other legal
    if(row$Other.Legal.Issue.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      fld = row$Field.of.Medicine.1
      Mention  = "Other Legal Issues"
      new_df <- rbind(new_df,c(yr,fld, Mention))
    }
    
}

```


```{r  , echo= FALSE}
#remove na's from data
new_df = new_df[-1,]
new_df= na.omit(new_df)

#count Frequencies
tot_df = plyr::count(new_df, c("yr", "field","Men"))

```

## Bubble plots and plots broken down by year

```{r , echo= FALSE }
# plt Legal issues mentioned by year (2021 values not prorated)
tot_df2 = plyr::count(new_df, c("yr", "Men")) 
tot = plyr::count(new_df, "yr")


x = ggplot(tot_df2, aes(x=yr, y=Men, size=freq, fill=freq)) +
    geom_point(alpha=0.7, shape=21,color= "white") +
    scale_size(range = c(.5, 16), name="freq") +
    scale_fill_viridis(discrete=FALSE, guide=FALSE, option="mako") +
    theme_classic() +
    theme(legend.position="bottom") +
    ylab("Legal Issue Mentioned") +
    xlab("Year") +
    ggtitle("Legal Issues Mentioned by Year") +
    theme(legend.position = "none") +
    theme(plot.title = element_text(hjust = 0.5))

print(x)


```

### Legal issue by year 2021 prorated assuming equal dist of papers published


```{r, echo= FALSE}
#prorate the 2021 values for after June
tot_dfpro = tot_df2
tot_dfpro$freq = ifelse(tot_df2$yr == 2021,
                       tot_df2$freq*2,
                       tot_df2$freq
                       )

#plot Prorated Bubble plot
x = ggplot(tot_dfpro, aes(x=yr, y=Men, size=freq, fill=freq)) +
    geom_point(alpha=0.7, shape=21,color= "white") +
    scale_size(range = c(.5, 15), name="freq") +
    scale_fill_viridis(discrete=FALSE, guide=FALSE, option="mako") +
    theme_classic() +
    theme(legend.position="bottom") +
    ylab("Legal Issue Discussed") +
    xlab("Year") +
    ggtitle("Frequency of Discussion of Legal Issues - by Year") +
    labs(caption= "*2021 values prorated")+
    theme(plot.caption = element_text(size = 6, hjust = 0))+
    theme(legend.position = "none") +
    theme(plot.title = element_text(hjust = 0))+
    theme(plot.title = element_text(size = 15,face = "bold"))

print(x)
```


```{r, echo= FALSE}
#plot prorated bar graph
q=ggplot(tot_dfpro, aes(fill=Men, y=freq, x=yr)) +
  scale_y_continuous(expand = expansion(mult = c(0, .1)))+
     geom_bar(position="stack", stat="identity")+
     theme(legend.position = "right") + 
     theme(plot.caption = element_text(size = 6, hjust = 0))+
     labs(x = "Year", y = "Frequency",fill = "Legal Issue Discussed", title = "Frequency of Discussion of Legal Issues - by Year",
          caption= "*2021 values prorated")
  
print(q)


```

## Line graph

### Projected assuming even dist of papers per year

```{r}
#Set Seed
set.seed(13)

#Split data into before and after prorated year
tot_dfcut = subset(tot_df2, yr!= 2021)
tot_dfprorate = tot_df2
tot_dfprorate$freq = ifelse(tot_df2$yr == 2021,
                       tot_df2$freq*2,
                       tot_df2$freq)

# create a vector of colors for line graph
col_vector = c("#b3de69", "#cab2d6","#546450", "#d1bb66", "#fc8d62", "purple", "brown", "pink", "#92d2f3", "#386cb0", "gold", "black")
# dummy labels as placeholder for first part of line graph
labs1 = c(c("label1", "label2", "label3", "label4","label5", "label6", "label7", "label8","label9", "label10", "label11", "label12"))

#Plot line Graph
plot = ggplot() + 
       #plot of current values
  geom_line(data = tot_dfcut,aes(x = yr, y = freq,colour = fct_reorder2(Men, yr, freq), group = Men),size = 1)+
  geom_point()+
  labs(x = "Year", y = "Frequency", color = "Legal Issue Discussed") +  
  labs(caption= "*2021 values prorated")+
  scale_color_manual(values= sample(col_vector, 12),labels = labs1)+ 
  theme_minimal() +
  ggtitle("Legal Issues Mentioned by Year") +
  theme(plot.title = element_text(hjust = 0.5))    +
  #add line for Prorated Values
  geom_line(data = tot_dfprorate,aes(x = yr, y = freq,colour = fct_reorder2(Men, yr, freq), group = Men),size   =    1,lty = 2)+
  labs(x = "Year", y = "Frequency", color = "Legal Issue Discussed") +  
  scale_color_manual(values= sample(col_vector, 12) )+ 
  ggtitle("Frequency of Discussion of Legal Issues - by Year") +
  theme(plot.title = element_text(hjust = 0))


plot

```



```{r}
set.seed(13)

#Create second version of plot above
plot <- ggplot() +
  geom_line(
    data = tot_dfcut,
    aes(x = yr, y = freq, colour = fct_reorder2(Men, yr, freq), group = Men),
    size = 1
  ) +
  geom_point() +
  labs(x = "Year", y = "Frequency", color = "Legal Issue Discussed") +
  labs(caption = "*2021 values prorated") +
  scale_color_manual(values = sample(col_vector, 12), labels = labs1) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),  # Remove major gridlines
    panel.grid.minor = element_blank(),  # Remove minor gridlines
    panel.border = element_blank(),  # Remove panel borders
    axis.line = element_line(colour = "black")  # Set axis lines
  ) +
  ggtitle("Legal Issues Mentioned by Year") +
  geom_line(
    data = tot_dfprorate,
    aes(x = yr, y = freq, colour = fct_reorder2(Men, yr, freq), group = Men),
    size = 1,
    lty = 2
  ) +
  labs(x = "Year", y = "Frequency", color = "Legal Issue Discussed") +
  scale_color_manual(values = sample(col_vector, 12)) +
  ggtitle("Frequency of Discussion of Legal Issues - by Year") +
  theme(plot.title = element_text(hjust = 0))+
  scale_y_continuous(expand = c(0, 0), limits = c(0, NA))+ labs(caption = c("*2021 Values Prorated", 
  expression(bold("Figure 2")))) + 
  theme(plot.caption = element_text(hjust=c(1, 0)))

plot
```


### Proated fom per year

```{r , echo= FALSE}
tot_df4 = plyr::count(new_df, c("yr", "field")) 

pro_df4= tot_df4
pro_df4$freq = ifelse(tot_df4$yr == 2021,
                       tot_df4$freq*2,
                       tot_df4$freq
                       )

#create plot
x = ggplot(pro_df4, aes(x=yr, y=field, size=freq, fill=field)) +
    geom_point(alpha=0.7, shape=21,color= "black") +
    scale_size(range = c(3, 15), name="field") +
    scale_fill_viridis(discrete = T, guide=FALSE, option="magma") +
    theme_classic() +
    theme(legend.position="bottom") +
    ylab("Field of Medicine") +
    xlab("Year") +
    ggtitle("Field of Medicine by Year") +
    theme(legend.position = "none") +
    theme(plot.title = element_text(hjust = 0.5))

print(x)
```

## Bar Graphs by legal issue by Feild of Medicine

```{r, echo = FALSE}
#create for loop for plots
bar_df = plyr::count(new_df, c("Men","field"))
#group other and unclear
bar_df$field[bar_df$field == "Unclear (Flag for review"] = "Other"
bar_df$field[bar_df$field == "Other: To Specify"] = "Other"
#Create a list of all legal issues mentioned
mentions =c("Privacy","Cyber Security","Safety/Quality","Bias/Discrimination",
            "Liability","Informed Consent","Capacity","Intellectual Property","Efficiency of Regulation","AI and Legal Personhood","Access/Insurance","Other Legal Issues")

#for loop plot all Mentions
for(i in mentions){
  varis = subset(bar_df, Men == i )
  x = ggplot(varis, aes(x = freq, y = reorder(field,freq))) +
  geom_col(fill = "#0099f9") +
    scale_x_continuous(expand = expansion(mult = c(0, .1)))+
  labs(
      title = paste(i,"Mentioned in Papers"),
      subtitle = "Broken Down by Field of Medicine"
    )+
  ylab("") +
  xlab("Frequency") +
  theme(
    plot.title = element_text(size = 15),
    plot.subtitle = element_text(face = "bold", color = "#0099f9", size = 10)
  )
  print(x)
  }
```

###DATA BY DISIPLINE OF JOURNAL

```{r , echo= FALSE}
## create DF 
yr = c(1)
disp = c(1)
Men = c(1)
bar_df = data.frame(yr,disp,Men)

# Iterate through the original dataframe and appent to new df based on what legal issue is mentioned
for(i in 1:nrow(data)) {
    row <- data[i,]
    # create row if an article mentions Privacy
    if(row$Privacy.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Type.of.Journal
      Mention  = "Privacy"
      bar_df <- rbind(bar_df,c(yr,disp, Mention))
    }
    
    # create row if an article mentions Cyber Security
    if(row$Cybersecurity.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Type.of.Journal
      Mention  = "Cyber Security"
      bar_df <- rbind(bar_df,c(yr,disp, Mention))
    }

    # create row if an article mentions Safety/Quality
    if(row$Safety.Quality.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Type.of.Journal
      Mention  = "Safety/Quality"
      bar_df <- rbind(bar_df,c(yr,disp, Mention))
    }

    # create row if an article mentions Bias
    if(row$Bias.Discrimination.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Type.of.Journal
      Mention  = "Bias/Discrimination"
      bar_df <- rbind(bar_df,c(yr,disp, Mention))
    }

    # create row if an article mentions Liability
    if(row$Liability.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Type.of.Journal
      Mention  = "Liability"
      bar_df <- rbind(bar_df,c(yr,disp, Mention))
    }

    # create row if an article mentions Informed Consent
    if(row$Informed.Consent.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Type.of.Journal
      Mention  = "Informed Consent"
      bar_df <- rbind(bar_df,c(yr,disp, Mention))
    }

    # create row if an article mentions Capacity
    if(row$Capacity.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Type.of.Journal
      Mention  = "Capacity"
      bar_df <- rbind(bar_df,c(yr,disp, Mention))
    }

    # create row if an article mentions IP
    if(row$Intellectual.Property.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Type.of.Journal
      Mention  = "Intellectual Property"
      bar_df <- rbind(bar_df,c(yr,disp, Mention))
    }

    # create row if an article mentions EOR
    if(row$Efficiency.of.Regulation.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Type.of.Journal
      Mention  = "Efficiency of Regulation"
      bar_df <- rbind(bar_df,c(yr,disp, Mention))
    }

    # create row if an article mentions AI and legal
    if(row$AI.Personhood.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Type.of.Journal
      Mention  = "AI and Legal Personhood"
      bar_df <- rbind(bar_df,c(yr,disp, Mention))
    }

    # create row if an article mentions Access/Insurance
    if(row$Access.Insurance.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Type.of.Journal
      Mention  = "Access/Insurance"
      bar_df <- rbind(bar_df,c(yr,disp, Mention))
    }


    # create row if an article mentions other legal
    if(row$Other.Legal.Issue.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Type.of.Journal
      Mention  = "Other Legal Issues"
      bar_df <- rbind(bar_df,c(yr,disp, Mention))
    }
}
```


## BY country

```{r, echo= FALSE}
#create new DF
yr = c(1)
con = c(1)
Men = c(1)
new_df2 = data.frame(yr,con,Men)

# Itterate and append based on legal issue mentioned, include country
for(i in 1:nrow(data)) {
    row <- data[i,]
    # create row if an article mentions Privacy
    if(row$Privacy.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      con = row$Corresp..Author..Country.
      Mention  = "Privacy"
      new_df2 <- rbind(new_df2,c(yr,con, Mention))
      # datalist[[i]] = c(yr,con,Mention)
    }
    
    # create row if an article mentions Cyber Security
    if(row$Cybersecurity.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      con = row$Corresp..Author..Country.
      Mention  = "Cyber Security"
      new_df2 <- rbind(new_df2,c(yr,con, Mention))
    }

    # create row if an article mentions Safety/Quality
    if(row$Safety.Quality.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      con = row$Corresp..Author..Country.
      Mention  = "Safety/Quality"
      new_df2 <- rbind(new_df2,c(yr,con, Mention))
    }

    # create row if an article mentions Bias
    if(row$Bias.Discrimination.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      con = row$Corresp..Author..Country.
      Mention  = "Bias/Discrimination"
      new_df2 <- rbind(new_df2,c(yr,con, Mention))
    }

    # create row if an article mentions Liability
    if(row$Liability.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      con = row$Corresp..Author..Country.
      Mention  = "Liability"
      new_df2 <- rbind(new_df2,c(yr,con, Mention))
    }

    # create row if an article mentions Informed Consent
    if(row$Informed.Consent.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      con = row$Corresp..Author..Country.
      Mention  = "Informed Consent"
      new_df2 <- rbind(new_df2,c(yr,con, Mention))
    }

    # create row if an article mentions Capacity
    if(row$Capacity.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      con = row$Corresp..Author..Country.
      Mention  = "Capacity"
      new_df2 <- rbind(new_df2,c(yr,con, Mention))
    }

    # create row if an article mentions IP
    if(row$Intellectual.Property.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      con = row$Corresp..Author..Country.
      Mention  = "Intellectual Property"
      new_df2 <- rbind(new_df2,c(yr,con, Mention))
    }

    # create row if an article mentions EOR
    if(row$Efficiency.of.Regulation.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      con = row$Corresp..Author..Country.
      Mention  = "Efficiency of Regulation"
      new_df2 <- rbind(new_df2,c(yr,con, Mention))
    }

    # create row if an article mentions AI and legal
    if(row$AI.Personhood.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      con = row$Corresp..Author..Country.
      Mention  = "AI and Legal Personhood"
      new_df2 <- rbind(new_df2,c(yr,con, Mention))
    }

    # create row if an article mentions Access/Insurance
    if(row$Access.Insurance.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      con = row$Corresp..Author..Country.
      Mention  = "Access/Insurance"
      new_df2 <- rbind(new_df2,c(yr,con, Mention))
    }


    # create row if an article mentions other legal
    if(row$Other.Legal.Issue.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      con = row$Corresp..Author..Country.
      Mention  = "Other Legal Issues"
      new_df2 <- rbind(new_df2,c(yr,con, Mention))
    }
    
}

```

```{r}
#Heatmap scale adjusted 
mapdat2 = data.frame(data$Year.of.Publication,data$Corresp..Author..Country.)
mapdat2 = mapdat2[-1,]

#remove 0 and group unknowns and correrct spelling errors in data
mapdat2$data.Corresp..Author..Country.[mapdat2$data.Corresp..Author..Country. == 'China '] = "China"
mapdat2$data.Corresp..Author..Country.[mapdat2$con == ''] = "Unknown"
mapdat2$data.Corresp..Author..Country.[mapdat2$con == ' '] = "Unknown"
mapdat2$data.Corresp..Author..Country.[mapdat2$data.Corresp..Author..Country. == '0'] = "Unknown"
mapdat2$data.Corresp..Author..Country.[mapdat2$data.Corresp..Author..Country. == 'Vietman'] = "Vietnam"
mapdat2$data.Corresp..Author..Country.[mapdat2$data.Corresp..Author..Country. == 'Australi'] = "Australia"
mapdat2$data.Corresp..Author..Country.[mapdat2$data.Corresp..Author..Country. == 'Germany '] = "Germany"
mapdat2$data.Corresp..Author..Country.[mapdat2$data.Corresp..Author..Country. == 'Germnay'] = "Germany"
mapdat2$data.Corresp..Author..Country.[mapdat2$data.Corresp..Author..Country. == 'France '] = "France"
mapdat2$data.Corresp..Author..Country.[mapdat2$data.Corresp..Author..Country. == 'United States'] = "USA"
mapdat2$data.Corresp..Author..Country.[mapdat2$data.Corresp..Author..Country. == 'United Kingdom'] = "UK"
mapdat2$data.Corresp..Author..Country.[mapdat2$data.Corresp..Author..Country. == 'England'] = "UK"
mapdat2$data.Corresp..Author..Country.[mapdat2$data.Corresp..Author..Country. == 'Amsterdam'] = "Netherlands"
mapdat2$data.Corresp..Author..Country.[mapdat2$data.Corresp..Author..Country. == 'The Netherlands'] = "Netherlands"
mapdat2$data.Corresp..Author..Country.[mapdat2$data.Corresp..Author..Country. == 'Ialy'] = "Italy"

#count by country
tot_map2 = plyr::count(mapdat2, c("data.Corresp..Author..Country.")) 

#make graph of count
ggplot(tot_map2, aes(x = freq, y = reorder(data.Corresp..Author..Country.,freq))) +
  geom_col(fill = "#0099f9") +
  scale_x_continuous(expand = expansion(mult = c(0, .1)))+
  labs(
    title = "Total Papers Published",
    subtitle = "By Country" )+
    ylab("Country") +
    xlab("Frequency") +
    theme(
    plot.title = element_text(size = 15),
    plot.subtitle = element_text(face = "bold", color = "#0099f9", size = 10)
  )

```

## By Continent

```{r, echo= FALSE}
#Add continent variable
mapdat2$conti = countrycode(sourcevar = mapdat2[, "data.Corresp..Author..Country."],
                            origin = "country.name",
                            destination = "continent")
mapdat2$conti[is.na(mapdat2$conti)] = "Unknown"

#Split apart the Americas
mapdat2[mapdat2$data.Corresp..Author..Country. =="Argentina" | mapdat2$data.Corresp..Author..Country. =="Brazil" , "conti"] <- "South America"
mapdat2[mapdat2$conti == "Americas" , "conti"] <- "North America"
#Count Per Continent
tot_map4 = plyr::count(mapdat2, c("conti"))

#Plot By Continent
ggplot(tot_map4, aes(x = freq, y = reorder(conti,freq))) +
  geom_col(fill = "#0099f9") +
  scale_x_continuous(expand = expansion(mult = c(0, .1)))+
  labs(
    title = "Total Papers Published",
    subtitle = "By Continent")+
    ylab("Continent") +
    xlab("Frequency") +
      theme(
    plot.title = element_text(size = 15),
    plot.subtitle = element_text(face = "bold", color = "#0099f9", size = 10))

  
```
## Graph USA removed

```{r, echo=FALSE}
#heatmap updated scales 
library(maps)
library(ggplot2)
library(countrycode)

tot_map2$data.Corresp..Author..Country.[tot_map2$data.Corresp..Author..Country. == 'United States'] = "USA"
tot_map2$data.Corresp..Author..Country.[tot_map2$data.Corresp..Author..Country. == 'United Kingdom of Great Britain and Northern Ireland'] = "UK"

#remove USA to visualize publication frequency without USA oversaturation
tot_map2$data.Corresp..Author..Country.[tot_map2$data.Corresp..Author..Country. == 'USA'] = "U"

#
world_map <- map_data("world")
world_map <- subset(world_map, region != "Antarctica")
x <- unique(world_map$region)
x = as.data.frame(x)
colnames(x)[1] ="con"
mapy = x
mapy$data.Corresp..Author..Country.= x$con
mapy= merge(x= mapy,y = tot_map2, by = "data.Corresp..Author..Country.", all.x = T)



#plot Heatmap without USA
ggplot(mapy) +
  geom_map(dat = world_map, map = world_map, aes(map_id = region, color = "DADADA"),
           fill = "white", color = "#7f7f7f", size = 0.25) +
  theme_minimal()+
  geom_map(map = world_map, aes(map_id = data.Corresp..Author..Country., fill = freq), size = 0.25) +
  scale_fill_viridis_c(option = "C",name = "Publications", direction = -1,na.value = "#DADADA") +
  expand_limits(x = world_map$long, y = world_map$lat) +
  labs(title = "Distribution of Papers Published (USA Removed)",
       subtitle = "Between 2012 -2021")+
  theme(plot.title = element_text(size = 15, face = "bold"),
        plot.subtitle = element_text(size = 9),
        axis.text.x = element_blank(),
        axis.text.y = element_blank())+
  ylab("") +
  xlab("")

```

## Graph USA included

```{r, echo=FALSE}
tot_map2$data.Corresp..Author..Country.[tot_map2$data.Corresp..Author..Country. == 'United States'] = "USA"
tot_map2$data.Corresp..Author..Country.[tot_map2$data.Corresp..Author..Country. == 'United Kingdom of Great Britain and Northern Ireland'] = "UK"
#readd USA
tot_map2$data.Corresp..Author..Country.[tot_map2$data.Corresp..Author..Country. == 'U'] = "USA"

world_map <- map_data("world")
world_map <- subset(world_map, region != "Antarctica")

x <- unique(world_map$region)
x = as.data.frame(x)
colnames(x)[1] ="con"
mapy = x
mapy$data.Corresp..Author..Country.= x$con
mapy= merge(x= mapy,y = tot_map2, by = "data.Corresp..Author..Country.", all.x = T)



#plot Heatmap With USA included
ggplot(mapy) +
  geom_map(dat = world_map, map = world_map, aes(map_id = region, color = "DADADA"),
           fill = "white", color = "#7f7f7f", size = 0.25) +
  theme_minimal()+
  geom_map(map = world_map, aes(map_id = data.Corresp..Author..Country., fill = freq), size = 0.25) +
  scale_fill_viridis_c(option = "C",name = "Publications", direction = -1,na.value = "#DADADA") +
  expand_limits(x = world_map$long, y = world_map$lat) +
  labs(title = "Distribution of Papers Published",
       subtitle = "Between 2012 -2021")+
  theme(plot.title = element_text(size = 15, face = "bold"),
        plot.subtitle = element_text(size = 9),
        axis.text.x = element_blank(),
        axis.text.y = element_blank())+
  ylab("") +
  xlab("")

```

```{r , echo= FALSE}
#new_df2 = new_df2[-1,]


#remove 0 and group unknowns
new_df2$con[new_df2$con == 'China '] = "China"
new_df2$con[new_df2$con == ''] = "Unknown"
new_df2$con[new_df2$con == '0'] = "Unknown"
new_df2$con[new_df2$con == 'Vietman'] = "Vietnam"
new_df2$con[new_df2$con == 'Australi'] = "Australia"
new_df2$con[new_df2$con == 'Germany '] = "Germany"
new_df2$con[new_df2$con == 'Germnay'] = "Germany"
new_df2$con[new_df2$con == 'France '] = "France"
new_df2$con[new_df2$con == 'United States'] = "USA"
new_df2$con[new_df2$con == 'United Kingdom'] = "UK"
new_df2$con[new_df2$con == 'England'] = "UK"
new_df2$con[new_df2$con == 'Amsterdam'] = "Netherlands"
mapdat2$data.Corresp..Author..Country.[mapdat2$data.Corresp..Author..Country. == 'The Netherlands'] = "Netherlands"

map_selected = mapdat2[,c("data.Corresp..Author..Country.","conti")]
new_df2 = merge(new_df2,map_selected,by.x = "con",by.y = "data.Corresp..Author..Country.")

```


## Discipline is Column I

```{r , echo= FALSE}
#create new data frame to transform data to count disciplines and solutions
yr = c(1)
displ = c(1)
sol = c(1)
new_df7 = data.frame(yr,displ,sol)
# iterate though data ad count the sols and disciplines
for(i in 1:nrow(data)){
      row <- data[i,]
      yr = row$Year.of.Publication
      displ  = row$Corresp..Author..Host.Faculty.
      sol = row$Solution.1
      sol2 = row$Solution.2
      sol3= row$Solution.3
      new_df7 <- rbind(new_df7,c(yr,displ, sol))
      if(sol2 != sol){
        new_df7 <- rbind(new_df7,c(yr,displ, sol2))
      }
      if(sol3 != sol & sol3 != sol2){
        new_df7 <- rbind(new_df7,c(yr,displ, sol3))
      }
} 

#merge others and shorten texts of solutions
new_df7 = new_df7[-1,]
new_df7$sol[new_df7$sol == "Other (Specify)"] = "Other"
new_df7$sol[new_df7$sol == "Other (specify)"] = "Other"
new_df7$sol[new_df7$sol == "0"] = "Other"
new_df7$sol[new_df7$sol == ""] = "Other"
new_df7$displ[new_df7$displ == "Unknown"] = "Unknown"
new_df7$displ[new_df7$displ == "Other health professions (Dentistry, nutrition, etc.)"] = "Other Health Professions"
sols = plyr::count(new_df7, c("displ","sol"))
sols2= sols
sols = subset(sols2, sol!= "N/A")
sols$sol[sols$sol == "Calls for Voluntary Improvements (By the Industry as a Whole, Individual Firms, Innovators, Providers, Etc.)" ] = "Voluntary Improvements"
sols$sol[sols$sol == "Mandatory Professional Training (As Part of Degree or Licencing Requirements or Continuing Education)" ] = "Mandatory Training"
sols$sol[sols$sol == "New Legislation or Regulations (E.g., A New Legislative Framework for Health-Related A.I.)"  ] = "New Legislation"
sols$sol[sols$sol == "Non-Binding Guidelines Issued By Others Bodies Without Regulatory Powers (E.g., The World Health Organization)" ] = "Non-Binding Guidelines"
sols$sol[sols$sol == "Other (Specify)"] = "Other"
sols$sol[sols$sol == "Professional Guidelines Issued by a Body with Regulatory Power (E.g., the College of Physicians and Surgeons or Professional Engineers Ontario, the F.D.A., Health Canada)" ] = "Professional Guidelines"
sols$sol[sols$sol == "Calls for Voluntary Improvements (By the Industry as a Whole, Individual Firms, Innovators, Providers, Etc.)" ] = "Voluntary Improvements"
sols$sol[sols$sol == "Reform of Existing Laws or Regulations (E.g., Amendments to Existing Privacy Laws)"] = "Reform of Existing Laws"

```


## Stacked bar graph discipline = col I

```{r, echo= FALSE}

sols$total1 <- ave(sols$freq, sols$sol, FUN = sum)
sols <- sols[order(sols$total1, decreasing = TRUE),]


#stacked bar graph of solutions by discipline
q=ggplot(sols, aes(fill=displ, y=freq, x=reorder(sol,-total1))) +
     geom_bar(position="stack", stat="identity") +
     scale_fill_brewer(palette  = "Spectral")
print(q + theme(legend.position = "right") +
        labs(x = "", y = "Frequency",fill = "Discipline") +
        scale_y_continuous(expand = expansion(mult = c(0, .1)))+
        #theme_minimal()+ 
        labs(
        title = paste("Proposed Solutions"),
        subtitle = "Broken Down by Discipline"

        )+
        theme(plot.title = element_text(size = 15,face = "bold"))+
        theme(axis.text.x = element_text(angle = -45, hjust=-.015)))

```

```{r, echo= FALSE}
sols$total <- ave(sols$freq, sols$displ, FUN = sum)
sols <- sols[order(sols$total, decreasing = TRUE),]


#plot discipline broken down by solutions
q=ggplot(sols, aes(fill=sol, y=freq, x=reorder(displ,-total))) +
     geom_bar(position="stack", stat="identity") +
     scale_fill_brewer(palette  = "Spectral")

print(q + theme(legend.position = "right") +
        labs(x = "", y = "Frequency",fill = "Solution") +
        scale_y_continuous(expand = expansion(mult = c(0, .1)))+
        labs(
        title = paste("Discipline Broken Down by Proposed Solutions")

        )+
        theme(plot.title = element_text(size = 15,face = "bold"))+
        theme(axis.text.x = element_text(angle = -45, hjust=-.015)))

```


## solutions by discipline of author bubble

```{r, echo= FALSE}

#bubble plot for displ against solution
x = ggplot(sols, aes(x=displ, y=sol, size=freq, fill=freq)) +
    geom_point(alpha=0.7, shape=21,color= "white") +
    scale_size(range = c(1.5, 10), name="freq") +
    scale_fill_viridis(discrete=FALSE, guide=FALSE, option="mako") +
    theme_classic() +
    theme(legend.position="bottom") +
    ylab("") +
    xlab("") +
    ggtitle("Solutions Proposed by Discipline") +
    theme(legend.position = "none") +
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.text.x = element_text(angle = 45, hjust=1))


print(x)
```



```{r, echo= FALSE}
## create DF 
yr = c(1)
disp = c(1)
Men = c(1)
bar_displ = data.frame(yr,disp,Men)


for(i in 1:nrow(data)) {
    row <- data[i,]
    # create row if an article mentions Privacy
    if(row$Privacy.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Corresp..Author..Host.Faculty.
      Mention  = "Privacy"
      bar_displ <- rbind(bar_displ,c(yr,disp, Mention))
      # datalist[[i]] = c(yr,disp,Mention)
    }
    
    # create row if an article mentions Cyber Security
    if(row$Cybersecurity.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Corresp..Author..Host.Faculty.
      Mention  = "Cyber Security"
      bar_displ <- rbind(bar_displ,c(yr,disp, Mention))
    }

    # create row if an article mentions Safety/Quality
    if(row$Safety.Quality.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Corresp..Author..Host.Faculty.
      Mention  = "Safety/Quality"
      bar_displ <- rbind(bar_displ,c(yr,disp, Mention))
    }

    # create row if an article mentions Bias
    if(row$Bias.Discrimination.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Corresp..Author..Host.Faculty.
      Mention  = "Bias/Discrimination"
      bar_displ <- rbind(bar_displ,c(yr,disp, Mention))
    }

    # create row if an article mentions Liability
    if(row$Liability.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Corresp..Author..Host.Faculty.
      Mention  = "Liability"
      bar_displ <- rbind(bar_displ,c(yr,disp, Mention))
    }

    # create row if an article mentions Informed Consent
    if(row$Informed.Consent.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Corresp..Author..Host.Faculty.
      Mention  = "Informed Consent"
      bar_displ <- rbind(bar_displ,c(yr,disp, Mention))
    }

    # create row if an article mentions Capacity
    if(row$Capacity.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Corresp..Author..Host.Faculty.
      Mention  = "Capacity"
      bar_displ <- rbind(bar_displ,c(yr,disp, Mention))
    }

    # create row if an article mentions IP
    if(row$Intellectual.Property.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Corresp..Author..Host.Faculty.
      Mention  = "Intellectual Property"
      bar_displ <- rbind(bar_displ,c(yr,disp, Mention))
    }

    # create row if an article mentions EOR
    if(row$Efficiency.of.Regulation.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Corresp..Author..Host.Faculty.
      Mention  = "Efficiency of Regulation"
      bar_displ <- rbind(bar_displ,c(yr,disp, Mention))
    }

    # create row if an article mentions AI and legal
    if(row$AI.Personhood.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Corresp..Author..Host.Faculty.
      Mention  = "AI and Legal Personhood"
      bar_displ <- rbind(bar_displ,c(yr,disp, Mention))
    }

    # create row if an article mentions Access/Insurance
    if(row$Access.Insurance.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Corresp..Author..Host.Faculty.
      Mention  = "Access/Insurance"
      bar_displ <- rbind(bar_displ,c(yr,disp, Mention))
    }


    # create row if an article mentions other legal
    if(row$Other.Legal.Issue.Mentioned == "Yes"){
      yr = row$Year.of.Publication
      disp = row$Corresp..Author..Host.Faculty.
      Mention  = "Other Legal Issues"
      bar_displ <- rbind(bar_displ,c(yr,disp, Mention))
    }
}

```

```{r, echo= FALSE}
#remove first row and correct spelling
bar_displ = bar_displ[-1,]
bar_displ$disp[bar_displ$disp == 'Other health professions (Dentistry, nutrition, etc.)'] = "Other Health Professions"
bar_displ$disp[bar_displ$disp == 'other'] = "Other"
bar_displ$disp[bar_displ$disp == 'Computer science'] = "Computer Science"
bar_displ$disp[bar_displ$disp == "Public health"] = "Public Health"
bar_displ2 = plyr::count(bar_displ, c("Men","disp")) 
```

```{r, echo= FALSE}
# Standardized text
co = data.frame(data$Corresp..Author..Host.Faculty.)
co$data.Corresp..Author..Host.Faculty.[co$data.Corresp..Author..Host.Faculty. == 'other'] = "Other"
co$data.Corresp..Author..Host.Faculty.[co$data.Corresp..Author..Host.Faculty.== 'Computer science'] = "Computer Science"
co$data.Corresp..Author..Host.Faculty.[co$data.Corresp..Author..Host.Faculty. == "Public health"] = "Public Health"
counts = plyr::count(co) 
total = sum(counts$freq)
counts[nrow(counts) + 1,] = c("Total",total)

```


```{r, echo= FALSE}
#for loop plot all legal issues mentioned broken down by Discipline
mentions =c("Privacy","Cyber Security","Safety/Quality","Bias/Discrimination",
            "Liability","Informed Consent","Capacity","Intellectual Property","Efficiency of Regulation","AI and Legal Personhood","Access/Insurance","Other Legal Issues") 


for(i in mentions){
  varis = subset(bar_displ2, Men == i )
  x = ggplot(varis, aes(x = freq, y =reorder(disp,freq))) +
  geom_col(fill = "#0099f9") +
    scale_x_continuous(expand = expansion(mult = c(0, .1)))+
  # theme_classic()+
  labs(
    title = paste(i,"Mentioned in Papers"),
    subtitle = "Broken Down by Discipline"
    
  )+
    ylab("Discipline") +
    xlab("Frequency") +
      theme(
    plot.title = element_text(size = 15),
    plot.subtitle = element_text(face = "bold", color = "#0099f9", size = 10)
  )
  print(x)
} 
```

##discipline broken down by legal issues discussed


```{r, echo = FALSE}
displs = unique(bar_displ2$disp)

#plot disipline broken dwn by legal issues raised
for(i in displs){
  varis = subset(bar_displ2, disp == i )
  x = ggplot(varis, aes(x = freq, y =reorder(Men,freq))) +
  geom_col(fill = "#0099f9") +
    scale_x_continuous(expand = expansion(mult = c(0, .1)))+
  # theme_classic()+
  labs(
    title = paste(i,"Papers"),
    subtitle = "Broken Down by Legal Issue Raised"
    
  )+
    ylab("Legal Issue Raised") +
    xlab("Frequency") +
      theme(
    plot.title = element_text(size = 15),
    plot.subtitle = element_text(face = "bold", color = "#0099f9", size = 10)
  )
  print(x)
} 

```



```{r, echo=FALSE}
#bubble chart of freq disp
x = ggplot(bar_displ2, aes(x=disp, y=Men, size=freq, fill=freq)) +
    geom_point(alpha=0.7, shape=21,color= "white") +
    scale_size(range = c(.5, 14), name="freq") +
    scale_fill_viridis(discrete=FALSE, guide=FALSE, option="") +
    theme_classic() +
    theme(legend.position="bottom") +
    ylab("Legal Issue Discussed") +
    xlab("Discipline") +
    ggtitle("Frequency of Discussion of Legal Issues") +
    labs(subtitle = "Across Disciplines")+
    theme(plot.caption = element_text(size = 6, hjust = 0))+
    theme(legend.position = "none") +
    theme(plot.title = element_text(hjust = 0))+
    theme(plot.title = element_text(size = 15,face = "bold"))+
    theme(axis.text.x = element_text(angle = 45, hjust=1))

print(x)
```



```{r}
#count occurances of country
amount = as.data.frame(table(data$Corresp..Author..Country.))
```


```{r}
#get solution values
solutions = sols[c("sol","total1")]
solutions = solutions[!duplicated(solutions),]
```

